---
title: "Homework 2"
subtitle: "MA 590 Special Topics: Causal Inference"
author: "Aukkawut Ammartayakun"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
    pdf_document:
        includes:
            in_header: "preamble.tex"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 1
## A
In a completely randomized experiment in which $n_1 = n_0 = n/2$, say $S_t^2 = S_c^2$ (the sets
of potential outcomes have the same variance in the experimental sample). What is the
smallest possible value for the true sampling variance of the difference in means estimator ($\mathbb{V}_{fp}(\hat{\tau}^{diff})$)? Justify your answer (with math)

(*hint: $S^2_{tc} = S^2_{t} + S^2_{c} - 2\rho S_tS_c$, where $\rho = \text{corr}\left\{Y(1),Y(0)\right\}$, the sample correlation of potential outcomes, $S_t = \sqrt{S_t^2}$, and $S_c = \sqrt{S_c^2}$, and recall that $-1 \leq \rho \leq 1$*)

### Solution

Given the expression here, we can say that $\mathbb{V}_{fp}(\hat{\tau}^{diff}) \leq \frac{2S_c^2 + 2S_t^2}{n}$
In another word, $\mathbb{V}_{fp}(\hat{\tau}^{diff}) \leq \frac{2S^2_{tc} + 2\rho S_tS_c}{n}$. Since $n,S_t,S_c$ and $S_{tc}$ are stictly positive, 
the minimum of maximum value of $\mathbb{V}_{fp}(\hat{\tau}^{diff})$ is $\frac{2S^2_{tc} - 2S_tS_c}{n}$ (i.e., negative correlation between two outcome). However, if there is no difference or constant treatment in two group. Namely, $y_i(0) = y_i(1), \forall i$, then $\mathbb{V}_{fp}(\hat{\tau}^{diff}) = 0$.

## B
With $n = 6$ and binary $Y$ , write out a table of potential outcomes for which $\mathbb{V}_{fp}(\hat{\tau}^{diff})$
attains its lowest value. What is the ATE?

### Solution

```{r table1, echo=FALSE}
df <- data.frame (y_1 = c(1,1,1,1,1,1), y_0 = c(0,0,0,0,0,0), Z_i = c(1,1,1,0,0,0))
df
```
ATE in this case would be $\bar{\tau} = 1-1 = 0$.

# Problem 2
In a Bernoulli randomized trial, each subject is randomized between $Z = 1$ and $Z = 0$
independently. Hence, $n_0$ and $n_1$ are random variables, and since $Pr(n_0 = 0) > 0$ and $Pr(n_1 = 0) > 0$, the difference-in-means estimator is not always defined. (The total
sample size n remains fixed, not random.) If $\mathbb{E}[Z_i] = Pr(Z_i = 1) = p$ for every subject $i$, show that the following is an unbiased
estimator for  $\bar{\tau}$:
$$\hat{\tau}_{IPW} = \frac{1}{n}\sum_{i=1}^n \left\{\frac{Z_iY_i}{p} - \frac{(1-Z_i)Y_i}{1-p}\right\}$$
(IPW stands for “inverse probability weighted” and will become important when we talk
about propensity scores.)

### Solution

We can see that if $Y_i = Z_iy_i(1) + (1-Z_i)y_i(0)$. Then,
$$\hat{\tau}_{IPW} =\frac{1}{n}\sum_{i=1}^n \left\{\frac{Z_iy_i(1)}{p} - \frac{(1-Z_i)y_i(0)}{1-p}\right\} $$
Taking the expectation, we can see that
$$\mathbb{E}[\hat{\tau}_{IPW}] =\frac{1}{n}\sum_{i=1}^n \left\{\frac{py_i(1)}{p} - \frac{(1-p)y_i(0)}{1-p}\right\}$$
which turns out to be the ATE or $\bar{\tau}$. Thus, conclude the unbiasness of the estimator.

# Problem 3
Had the Lady Tasing Tea experiment been a Bernoulli randomized trial, so that each cup
would be randomized to tea-first or milk-first independently, what would be the $p$-value
(using Fisher’s method) had the lady gotten all 8 cups right?

### Solution

Let assume that we don't care whether the trial we have will contains both tea-first and milk-first. Then, there are $2^8 = 256$ possible outcomes. Even so, there is only one way to orient the test to match with the given answer. The probability of getting all 8 cups to align with the guessing is $\frac{1}{256}$. Thus, the $p$-value is $\frac{1}{256}$.

